Network Configuration Synthesis with Abstract Topologies

### 2016-11-09 paper/abstraction/appendix.tex
update pgir images. describe concretization functions

  \begin{minipage}[t]{.46\linewidth}
  \hdr{ABGP Syntax}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rclr}
     abgp &::=& l_1 \rightarrow {rc}_1, ~\dots,~ l_k \rightarrow {rc}_k & \textit{abgp policy} \\
     rc   &::=& t_1 \rightarrow {pc}_1, ~\dots,~ t_k \rightarrow {pc}_k & \textit{router config} \\
     pc   &::=& ma_1, \dots, ma_k & \textit{prefix config} \\
     ma   &::=& n_1 : ({ns}_1, n_2) \rightarrow ({ns}_2, n_3) & \textit{match action} \\
     ps   &::=& \{ t_1, ~\dots,~ t_k \} & \textit{predicates} \\
     ns   &::=& \{ l_1, ~\dots,~ l_k \} & \textit{peers} \\
  \end{array} \]%

>>>>

  \begin{minipage}[t]{.46\linewidth}
  \hdr{ABGP Syntax}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rclr}
     abgp &::=& l_1 \rightarrow {rc}_1, ~\dots,~ l_k \rightarrow {rc}_k & \textit{abgp policy} \\
     rc   &::=& t_1 \rightarrow {pc}_1, ~\dots,~ t_k \rightarrow {pc}_k & \textit{router config} \\
     pc   &::=& ma_1, \dots, ma_k & \textit{prefix config} \\
     ma   &::=& n_1 : ({ns}_1, n_2) \rightarrow ({ns}_2, n_3) & \textit{match action} \\
     ns   &::=& \{ l_1, ~\dots,~ l_k \} & \textit{peers} \\
     t    &::=& \$x \BNFALT d.d.d.d/[d..d] & \textit{predicate} \\
  \end{array} \]%

### 2016-11-09 paper/abstraction/paper.tex
update pgir images. describe concretization functions

\subsection{Compilation Correctness}

@@

  \begin{minipage}[t]{.46\linewidth}
  \hdr{ABGP Syntax}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rclr}
     abgp &::=& l_1 \rightarrow {rc}_1, ~\dots,~ l_k \rightarrow {rc}_k & \textit{abgp policy} \\
     rc   &::=& t_1 \rightarrow {pc}_1, ~\dots,~ t_k \rightarrow {pc}_k & \textit{router config} \\
     pc   &::=& ma_1, \dots, ma_k & \textit{prefix config} \\
     ma   &::=& n_1 : ({ns}_1, n_2) \rightarrow ({ns}_2, n_3) & \textit{match action} \\
     ps   &::=& \{ t_1, ~\dots,~ t_k \} & \textit{predicates} \\
     ns   &::=& \{ l_1, ~\dots,~ l_k \} & \textit{peers} \\
  \end{array} \]%

  \end{minipage}
  %
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{Compilation}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{l}
     \text{compile}_\text{ABGP}( (t_1,PG_1,ord_1), \dots, (t_k,PG_k,ord_k) ) = \\
     ~~~~~ \{ l \rightarrow rc ~\vert~ \\
     ~~~~~~~~~ l \in internal(V), \\
     ~~~~~~~~~ rc = \bigcup_i~ \{ t_i \rightarrow ma_1, \dots, ma_k ~\vert~ \\
     ~~~~~~~~~~~~~~~~~~~~~ m = (l,q_m) \in PG_i, \\
     ~~~~~~~~~~~~~~~~~~~~~ in = \{ b ~\vert~ (b,q_n) \in \text{adjIn}(PG_i,m) \} \\
     ~~~~~~~~~~~~~~~~~~~~~ out = \{ c ~\vert~ (c,\_) \in \text{adjOut}(PG_i,m) \} \\
     ~~~~~~~~~~~~~~~~~~~~~ ma_j = ord_i(m) : (in,q_n) \rightarrow (out,q_m) \} \} \\
     \\
     \text{compile}( p_1, \dots, p_k ) = \\
     ~~~~~~ \text{compile}_\text{ABGP}(\text{compile}_\text{PG}(p_1), \dots, \text{compile}_\text{PG}(p_k)) \\
  \end{array} \]%
  \end{minipage}

@@

An important observation is that the topology abstraction $f$ is preserved throughout compilation. This is reflected in the structure of the template and concrete configurations from Figure~\ref{fig:bgp-configs}. At a first cut, the configurations appear the same except for the fact that (1) the abstract locations are replaced with the union of all concrete locations in the concrete configurations, and (2) the prefix templates are replaced with an entry for each unique prefix in the concrete policy.

In this section, we formalize this observation. We introduce a simple language for BGP configurations called ABGP and define compilation over both concrete and abstract topologies to ABGP. In addition we define a concretization function for both \sysname policies as well as ABGP configurations that take an abstract policy/configuration and turn it into a concrete policy/configuration. The main theoretical result is that the compilation and concretization functions commute. This means we can perform compilation over the abstract representation to generate templates, and only concretize these templates when we need to instantiate the network.

@@

We introduct a simple language for BGP configurations called ABGP. Figure~\ref{fig:concretization} shows the syntax for an ABGP policy.

 An ABGP policy consists of a map from topology locations $l$ to router configurations $rc$. Each router configuration $rc$ is a sequence of prefix (templates) $t$ and their corresponding prefix configuration $pc$. A prefix configuration $pc$ is a sequence of match actions $ma$. Each match action $ma$ consists of a preference value $n_1$, followed by a set of imports from neighbors $ns_1$ and community value (\IE, state of the product graph) $n_2$, as well as export neighbors $ns_2$ and the updated community value $n_3$.

 \vspace{1em}

We assume a particular router $l$ appears at most once in any abgp specification.
We often build configs using set-builder notation.  For instance,
$\{l \rightarrow rc ~\vert~ l \in V \wedge p(l,rc)\}$ denotes the abgp policy

\[ \begin{array}{c}
  l_1 \rightarrow rc_{11},... \\
  ...           \\
  l_k \rightarrow rc_{1k},... \\
\end{array} \]

where $l_1, ..., l_k$ are the nodes in the topology
and each $rc_{ij}$ satisfies $p(l_i,rc_{ij})$.

@@

Figure~\ref{fig:abgp-syntax} also defines compilation from a \sysname policy to an ABGP policy. It starts by compiling separately each constraint of the form $p_i = (t_i ~\Path~ r_{i1} ~\Prefer~ r_{ik})$ to a product graph $PG = (G', s, P)$, the particular test $t_i$, and the preference ordering $ord_i$.

The function $compile_{ABGP}$ then takes tuples $(t_i, PG_i, ord_i)$ and returns an ABGP policy. It looks at each incoming neighbor with the same community state $q_n$ and matches any of those neighbor with the preference provided by the ordering $ord$ obtained from the search for local preferences. It then exports this advertisement to each adjacent peer along an outgoing edge and adds the community value for the current state $q_m$.

@@

We define what it means to concretize an abstract \sysname policy. 
Concretization takes as input, a policy and a context $\Gamma$ mapping from a template prefix to a set of pairs of a concrete prefix and a concrete location where the prefix is owned.

For each regular expression $r$, we replace each abstract location with the sum over its corresponding concrete locations under the inverse homomorphism $f^{-1}$. For each template variable, we replace it with a new entry for each of its corresponding concrete values under the supplied mapping $\Gamma$. We also intersect the resulting concretized regular expression with $\text{end}(l)$ to capture the fact that traffic is owned by the concrete location $l$. \textbf{Note}: we need the side condition that $\text{end}(l) \cap con(r_i,\Gamma) \neq \emptyset$ if $con(r_i,\Gamma) \neq \emptyset$.
We can also define concretization over the ABGP policy as:

We use the symbol $\bullet$ to denote that a particular value in a sequence is removed from the sequence.

\begin{thm}
  For any context $\Gamma$, concrete topology $G$, homomorphism $f : G \rightarrow G^A$, and policy $pol$,
  $$\Con(\text{compile}(pol), \Gamma,f,G) = \text{compile}(\Con(pol,\Gamma,f))$$
\end{thm}

@@

Suppose that we want to expand the concrete datacenter by adding an additional tor router in the TG role. Per the network routing policy, this new ToR router will advertise its own owned prefix. Because the new topology matches the abstraction, the compiled templates will remain the same. However, in the spine router configurations, the match on the global prefix template variable \$GP will be expanded to include the new prefix owned by the ToR. In other words, the concrete configurations obtained after substitution will include an additional match on the new prefix. Unfortunately, this means a small change to the topology can result in many configurations far from the change needing to change too.

The key observation is that the template variables tell us exactly where change can happen. For example, the only way in which the spine routers depend on anything nonlocal is in the prefix template variables. Instead, we replace each match on a template variable (\EG, matching \$GP) with a test on a new community tag that we associate with each template variable. So long as we ensure that routers always add the appropriate tags to the prefixes they originate, then the policies will be equivalent.

Specifically, this ensures that each router has a template that has no dependency on any other non-adjacent router. If we make a change to the concrete network that matches the abstraction, then the templates do not change, and thus only configurations adjacent to the changes must also change.

>>>>

\section{Concretization}

@@

  \begin{minipage}[t]{.46\linewidth}
  \hdr{ABGP Syntax}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rclr}
     abgp &::=& l_1 \rightarrow {rc}_1, ~\dots,~ l_k \rightarrow {rc}_k & \textit{abgp policy} \\
     rc   &::=& t_1 \rightarrow {pc}_1, ~\dots,~ t_k \rightarrow {pc}_k & \textit{router config} \\
     pc   &::=& ma_1, \dots, ma_k & \textit{prefix config} \\
     ma   &::=& d_1 : ({ns}_1, d_2) \rightarrow ({ns}_2, d_3) & \textit{match action} \\
     ns   &::=& \{ l_1, ~\dots,~ l_k \} & \textit{peers} \\
     t    &::=& \$x \BNFALT d.d.d.d/[d..d] & \textit{predicate} \\
  \end{array} \]%

  \end{minipage}
  %
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{Compilation}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{l}
     \text{compile}_\text{ABGP}( (t_1,PG_1,ord_1), \dots, (t_k,PG_k,ord_k), G ) = \\
     ~~~~~ \{ l \rightarrow rc ~\vert~ \\
     ~~~~~~~~~ l \in internal(G.V), \\
     ~~~~~~~~~ rc = \bigcup_i~ \{ t_i \rightarrow ma_1, \dots, ma_k ~\vert~ \\
     ~~~~~~~~~~~~~~~~~~~~~ m = (l,q_m) \in PG_i, \\
     ~~~~~~~~~~~~~~~~~~~~~ in = \{ b ~\vert~ (b,q_n) \in \text{adjIn}(PG_i,m) \} \\
     ~~~~~~~~~~~~~~~~~~~~~ out = \{ c ~\vert~ (c,\_) \in \text{adjOut}(PG_i,m) \} \\
     ~~~~~~~~~~~~~~~~~~~~~ ma_j = ord_i(m) : (in,q_n) \rightarrow (out,q_m) \} \} \\
     \\
     \text{compile}( p_1, \dots, p_k, G) = \\
     ~~~~~~ \text{compile}_\text{ABGP}(\text{compile}_\text{PG}(p_1,G), \dots, \text{compile}_\text{PG}(p_k,G)) \\
  \end{array} \]%
  \end{minipage}

@@

One interesting observation is that the spine template configuration and concrete spine configurations are structured similarly. In this section, we formalize this observation. We introduce a simple language for BGP configurations called ABGP and define compilation policies over both concrete and abstract topologies to ABGP. In addition we define a concretization function for that translates an abstract \sysname policy or ABGP configuration to a concrete \sysname policy or ABGP configuration respectively. The main result is that the compilation and concretization functions commute. This means performing compilation over the abstract representation first allows us to generate templates, and only concretize these templates when we need to instantiate the network.

@@

We introduct a simple language for BGP configurations called ABGP. Figure~\ref{fig:concretization} shows the syntax for an ABGP policy. An ABGP policy maps topology locations $l$ to router configurations $rc$. Each router configuration $rc$ is a sequence of predicates $t$ and their corresponding configuration $pc$. A prefix configuration $pc$ is a sequence of match actions. Each match action $ma$ consists of a preference value $d_1$, followed by a set of imports from neighbors $ns_1$ and community value (\IE, state of the product graph) $d_2$ expected from those neighbors, as well as export neighbors $ns_2$ and the community value $d_3$ to attach before sending to those neighbors.

@@

Figure~\ref{fig:concretization} defines compilation from a \sysname policy to an ABGP policy. It starts by compiling ($\text{compile}_\text{PG}$) separately each constraint of the form $p_i = (t_i ~\Path~ r_{i1} ~\Prefer~ r_{ik})$ to a product graph $PG = (G', s, P)$. The function $\text{compile}_\text{PG}$ returns the test $t_i$, product graph $PG_i$, and the preference ordering $ord_i$. 
%
The function $\text{compile}_\text{ABGP}$ then takes a sequence of tuples $(t_i, PG_i, ord_i)$ and returns an ABGP policy. For each internal router, it looks at each set of neighbor with the same community state $q_n$ and matches any of those neighbors with the preference provided by the ordering $ord$. It then exports this advertisement to each adjacent peer along an outgoing edge and adds the community value for the current state $q_m$.

%We assume a particular router $l$ appears at most once in any abgp specification.
We build configs using set-builder notation.  For instance,
$\{l \rightarrow rc ~\vert~ l \in V \wedge p(l,rc)\}$ denotes the abgp policy
%
\[ \begin{array}{c}
  l_1 \rightarrow rc_{1},... \\
  ...           \\
  l_k \rightarrow rc_{k},... \\
\end{array} \]

where $l_1, ..., l_k$ are the nodes in the topology
and each $rc_{i}$ satisfies $p(l_i,rc_{i})$. We also use the $\bigcup$ operator to denote sequence concatenation.

@@

Concretization for a \sysname policy takes as input, the policy and a context 
$\Gamma : Var \rightarrow 2^{Prefix \times V}$, 
which maps from a template prefix to a set of pairs of a concrete prefix and a concrete location that "owns" the prefix.
%
For each regular expression $r$, we replace each abstract location with the sum over its corresponding concrete locations under the inverse homomorphism $f^{-1}$. For each template variable, we replace it with a new entry for each of its corresponding concrete prefixes under the supplied mapping $\Gamma$. We also intersect the resulting concretized regular expression with $\text{end}(l)$ to capture the fact that this traffic is owned by the concrete location $l$.

We also define concretization over the ABGP policy in a similar manner. We replace each abstract router configuration $l \rightarrow rc$ a new entry for each concrete location $\ell \in f^{-1}(l)$. The concrete location $\ell$ is passed as a parameter to $con$ to remember the current router.
%
To concretize a prefix policy with a template ($\$x \rightarrow pc$) we replace it with a sequence of policies, one for each prefix in context ($(pfx,l) \in \Gamma(x)$). Match action statements where the set of import neighbors is the singleton set containing the product graph start node $\{ start \}$ and we are not currently at looking at the router configuration for the location specified in the context $l$, then we drop the match action. This is denoted with the $\bullet$ symbol. Finally, to concretize a set of neighbors, we again use $f^{-1}$ to find the set of concrete neighbors under the inverse homomorphism, but then restrict it to those that are actually connected in the underlying concrete topology with $G.E$.

\begin{thm}
  For any context $\Gamma$, topology $G$, abstract topology $G^A$, homomorphism $f : G \rightarrow G^A$, and policy $pol$,
  $$\Con(\text{compile}(pol,G^A), \Gamma,f,G) = \text{compile}(\Con(pol,\Gamma,f),G)$$
\end{thm}

@@

Suppose that we want to expand our example concrete datacenter by adding an additional tor router in the TG role. Per the network routing policy, this new tor router will advertise its new owned prefix. Because the new topology matches the abstraction, the compiled templates will remain the same. However, in the spine router configurations, the match on the global prefix template variable \$GP will be expanded to include the new prefix owned by the added tor. In other words, the concrete configurations obtained after substitution will include an additional match on the new prefix. Unfortunately, this means that a small change to the topology can result in many configurations far from the change needing to change too.

The key insight is that the template variables tell us exactly where change can happen. For example, the only way in which the spine routers depend on anything non-local is in the match on prefix template variables. Instead, we can associate a new unique community tag for each template variable like \$GP. We can replace each match on \$GP with a test on the new community tag. The new added tor will now simply attach the new community tag to advertisements for its prefix so downstream routers can just match the community tag. This ensures that each router has a template that has no dependency on any other non-adjacent router. If we make a change to the concrete network that matches the abstraction, then the templates do not change, and thus only configurations adjacent to the changes must also change.
